---
title: "Fishbase Queries"
author: "Cori Lopazanski"
date: "7/1/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(freeR) # where Chris Free's `fishbase` function lives
library(tidyverse)
library(rfishbase)
library(here)
```

***
This document retrieves information from FishBase (using `rfishbase` package) and cross-examines species lists from the RAM Legacy Database.

### Document Outline: 

- Attempt to use Chris Free's function `fishbase`, a wrapper for the `rfishbase` package that provides the information in cleaned tables
- Using `rfishbase` to retrieve information from FishBase ecology table
- Load the RAM legacy database and retrieve species/stock information
- Restrict the RAM species list for the FAO regions overlapping with mangrove geographic extent
- Cross-reference FishBase species list with the geographically-relevant RAM Legacy species list to see matches

***

#### A. Using `freeR::fishbase` 

Test out Chris's function: ?fishbase tells information about the arguments for the function. 

- Current focus is getting habitat information, which I know from searching the back end of FishBase is mainly located in the 'ecology' table
- Try to use 'fishbase' to get the ecology level information
- Error. Talked to Chris, he doesn't know what's up. Probably a package dependency?

```{r}
# Retrieve species names from the RAM database:
#ram.sp <- stock$scientificname

# Use that list to query FishBase?
#eco <- fishbase(dataset = "ecology", species = ram.sp, cleaned = T)
```


#### B. Using `rfishbase` to get information from FishBase

Link to fishbase Appendix: https://www.fishbase.in/FishOnLine/English/FOL_Appendix.htm

Use rfishbase to get the relevant tables:
```{r}
# Retrieve ecology table for all species 
fb.eco.table <- ecology() # Currently 34329 obs

# Retrieve species table for all species
fb.species.table <- species()

# Retrieve spawning table for all species
fb.spawning.table <- spawning() 

# Notes about other tables:
# fb.repro.table <- reproduction() - no relevant habitat information

```

Note of relevant fields in the tables after review:

- Species:
  - `Fresh`, `Brack`, and `Saltwater` - environments where the fish are found
- Ecology: various habitats listed under the ecology table
  - `Estuaries`
  - `Saltmarsh`
  - `Mangroves`
  - `MarshesSwamps`
  - `AddRems` - explanation of 'additional remarks' which often contain info about life history and ontogenetic shifts
- Spawning:
  - `SpawningGround` - listed as "coastal", "estuarine", or NA
  - `Spawningarea`- more specific spawning area (if any)


```{r}
# Create subsets of tables to make easier to work with
fb.sp.subset <- fb.species.table %>% 
  select("SpecCode", "Species", "Fresh", "Brack", "Saltwater") %>% 
  filter(Brack == -1) %>% # only keep brackish water species.... otherwise list is 18k 
  unique(.)
  
fb.eco.subset <- fb.eco.table %>% 
  select("SpecCode", "Species", "Estuaries", "Saltmarshes", "Mangroves", "MarshesSwamps", "AddRems") %>% 
  filter_at(vars(Estuaries:MarshesSwamps), any_vars(.!=0)) %>%  # only keep where one of the relevant columns is -1 (present) - 2013 obs.
  unique(.)

#fb.eco.dup <- fb.eco.subset[duplicated(fb.eco.subset$Species),] # There are some weird duplicates, not sure what's up there but prolly ok 
  
fb.spawning.subset <- fb.spawning.table %>% 
  select("SpecCode", "Species", "SpawningGround") %>% # chose not to include spawning area
  filter(SpawningGround %in% c("coastal", "estuarine")) %>%  #coastal and estuarine spawners 
  unique(.) %>% 
  mutate(spawner = -1) %>% 
  pivot_wider(names_from = SpawningGround, values_from = spawner) %>% 
  group_by(SpecCode, Species) %>% 
  summarize(Coastal_Spawner = sum(coastal, na.rm = T),
            Estuarine_Spawner = sum(estuarine, na.rm = T)) # That was a dumb way to do that but whatever
# Not including spawning area because multiple spawning areas results in multiple entries per species which is annoying


# Join the subsets together:
fb.ecosp <- full_join(fb.sp.subset, fb.eco.subset, by = c("SpecCode","Species"))
fb.all <- full_join(fb.ecosp, fb.spawning.subset, by = c("SpecCode", "Species")) #4102 obs

fb.all.duplicates <- fb.all[duplicated(fb.all$Species),] # The duplicates....
```

Looking at the datafarame `fb.all` - most of the 'coastal spawner' species don't have any additional habitat information, which probably won't really help. 

Make a subset of the `fb.all` dataframe (which contains the species + ecology + spawning table, subsetted) which excludes any row where all relevant habitat metrics other than 'SpawningGround' are NA
```{r}
fb.all.subset <- fb.all %>% 
  filter_at(vars(Fresh:MarshesSwamps), any_vars(!is.na(.))) # 3791 possible species miiiiight use mangroves
```

#### C. Load RAM Legacy Database and retrieve species list

Load the RAM database workspace, and save the species list:
```{r}
# Load the RAM legacy database workspace v4.491:
load("DBdata[asmt][v4.491].RData")

# The species list can easily be retrieved from the stock or metadata tables:
# Save the species list as ram.sp
ram.sp <- unique(stock$scientificname) # 361 species in the RAM database
```


#### D. Restrict RAM species list based on FAO regions overlapping with geographic extent of mangroves

Use method from Sheaves 2017 for narrowing down geographic extent of species list

- Filter out species that occur in FAO areas that encompass the geographic range of mangroves
- A fishery likely doesn't use mangroves if its distribution doesn't overlap with mangrove cover... obviously
- Sheaves' list of FAO regions: 01, 03, 04, 06, 31, 34, 41, 47, 51, 57, 61, 71, 77, 81, 87
- FAO Areas are listed in the RAM database under the 'stock' dataframe, `primary_FAOarea`

```{r}
# Save the mangrove-range FAO region number list
fao.mangrove.regions <- c(1, 3, 4, 6, 31, 34, 41, 47, 51, 57, 61, 71, 77, 81, 87)

# The column that has the FAO area number is a character and 
# causing problems with the numbers <10 so will convert that 
# column to numeric to be able to compare to the list above
numericstock <- stock %>% 
  mutate(primary_FAOarea = as.numeric(primary_FAOarea))

# Filter the RAM stocks list for stocks that are located 
# in the FAO regions that overlap with mangroves
ram.fao.subset <- numericstock %>% 
  filter(primary_FAOarea %in% fao.mangrove.regions) 

# 368 potential mangrove-associated stocks in RAM based on geography

# How many species make up those 368 stocks?
ram.mang.sp <- unique(ram.fao.subset$scientificname)

# 197 potential mangrove-associated species in RAM based on geography
```

***

#### **Summary So Far and Important Dataframes** 

We've now loaded information from both FishBase and RAM - 

- The FishBase subset (`fb.all.subset`) contains all of the species listed in the species, ecology, and spawning tables. The species list is made up of species that:
  - Use brackish water AND/OR have at least one relevant habitat characteristic (estuaries, mangroves, saltmarshes, marshes/swamps) 
  - Recall: this subset contains information about the spawning ground, but species were excluded if they did not have information about either water type or habitat
  - This is likely the most exhaustive list (species that could have possibly swam past a mangrove once in their life) - best strategy will likely be to narrow down the list based on specific identifiers
- The RAM subset (`ram.fao.subset`) contains all of the STOCKS listed in the RAM database that are located within FAO regions that overlap with the geographic extent of mangroves (as in Sheaves 2017)
  
***

#### Cross-reference the FishBase and RAM dataframes to check out potential species matches

```{r}
# Which of the species listed in the FishBase Subset are also in the RAM Subset?
fb.inram<- fb.all.subset %>% 
  filter(Species %in% ram.fao.subset$scientificname)
# 42 species in the FishBase Subset are in RAM FAO Subset
# This means: these species are listed in FB as either occupying brackish water AND/OR matching one of the estuarine habitat characteristics, have assessments in the RAM database, and are located in FAO Regions that overlap with geographic mangrove extent

# How many stocks for those 42 species?
fbram.stocks <- ram.fao.subset[which(ram.fao.subset$scientificname %in% fb.inram$Species),] #70 stocks


# What about ones that just say they use mangroves?
fb.mangroves <- fb.all.subset[which(fb.all.subset$Mangroves == -1),] # 609 species in fishbase subset that have mangroves listed as habitat
fb.mang.inram <- fb.mangroves[which(fb.mangroves$Species %in% ram.fao.subset$scientificname),] # 7 species in RAM use mangroves

# How many stocks make up those 7 species?
fbram.mang.stocks <- ram.fao.subset[which(ram.fao.subset$scientificname %in% fb.mang.inram$Species),] # 9 stocks

```

Mangroves species list:

* Australian salmon
* Blacktip shark
* Cobia
* Goliath grouper
* Largehead hairtail (China/Japan)
* Sea mullet (Australia)
* Mutton snapper <<< THIS ONE MIGHT BE GOOD!!


