---
title: "seaaroundus-fishbase-queries.Rmd"
author: "Kirby Bartlett"
date: "7/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#library(freeR) # where Chris Free's `fishbase` function lives
library(tidyverse)
#library(rfishbase)
library(here)
library(sf) # Stores information about the CRS and classes of the spatial data
library(raster) # Package that deals with raster algebra
library(rgdal) # Defining CRS

```


Set-Up for Downloading Data Using the Google Drive File Stream

# Create a Root to Extract Files 
if(Sys.info()["sysname"] == "Windows"){ #If operating system is "Windows" use G: Drive
  team_path <- "G:/"
} else { # If it is not (i.e. Mac), use the Google Drive File Stream
  team_path <- "/Volumes/GoogleDrive/.shortcut-targets-by-id/1GpvrcGzmRSROKwiSRhlx3wUH3WCLFULw/"
}

#Create the Full File Path
path <- paste0(team_path, "fishery-habitat-project/Data/Raw_Data/")



```{r}

# Read in downloaded Watson 2000-2004 industrial fishing data
watson2000.2004_ind <- read_csv(paste0(team_path, "CatchInd2000_2004.csv"))


# Read in the cell codes csv which includes lat/long info 
cell_codes <- read_csv("CellCodes.csv")

# Read in the index file 
index_2000 <- read_csv("IndexInd.csv") %>% 
  filter(Year = "2000")

# Join the spatial information with the cell codes for each location
joined_ind_2000_2004 <- full_join(watson2000.2004_ind, cell_codes, by = "Cell") %>% 
  select(LonCentre, LatCentre, Reported, IUU, Discards, Cell, ID) %>% 
  rename(Reported_ind = Reported,
         IUU_ind = IUU,
         Discards_ind = Discards)

# Join together the index information with the catch/cell code information. Make sure longitude and latitude are the first two columns for plotting purposes.
# Select for the columns that you want to be made into rasters
index_catch_join <- full_join(joined_ind_2000_2004, index_2000, by = "ID") %>% 
  select(LonCentre, LatCentre, Reported, IUUTotal, Discards, Reported_ind, IUU_ind)
  

# Raster from XYZ. Plot the lat/long information. It will make a raster for every column you have.
ind_xyz_2000_2004 <- rasterFromXYZ(index_catch_join, crs = "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs")

# Assign column names to layer names
names(ind_xyz_2000_2004) <- tail(colnames(index_catch_join), 3) 

```

```{r}
# Read in tif of mangroves in Indonesia that was exported from ArcPro
# Mangrove_2000 <- raster("a2000/a2000.tif")
```

